Report on Late Winter2013 Production: Image Differencing

### 2013-03-26 W13report.tex
Fleshed out WCS rms and offset sections.

We ran several pre--productions to tune the following: the dimensions
of the kernels and {\tt KernelCandidates}; the widths of the Gaussians
(sigGauss) based on the FWHM of the input images; and the scaling of
the Gaussians ($\beta$) in the basis sequence.
Figure~\ref{fig_galleryBad} provides a rogue's gallery of image
subtraction failures, centered on {\tt KernelCandidates}, and outlines
reasons for each failure.  Figure~\ref{fig_galleryGood} provides a
gallery showing what the results from a successful image subtraction
appear like, for postfiltering on the {\it top} and prefiltering on
teh {\it bottom}.

@@

We implemented two simple perturbations of the inputs to the
image--to--image RegisterTask: we first added a random offset to each
object's (x,y)-coordinates, with an amplitude that was specified in
the Config and multiplied by a random number pulled from a normal
distribution; and we added a DC offset to the coordinates at an
amplitude specified in the Config.  These offsets, added to the Source
coordinates, will affect misalignments of the objects in the
registered images, as the registration is done assuming the specified
positions are correct.  In this way we are able to investigate how
random uncertainties and bulk astrometric offsets impact the false
positive rate.  We explicitly do {\it not} investigate spatial
variation in these offsets, using for example a pincushion distortion.
We anticipate that this latter effect will be most important for
spatial interpolation and extrapolation of the matching kernel,
yielding a dipole residual field associated with the distortion.
However, the ability of the software to model out these spatial
distortions will certainly fail if the local kernel solutions (from
which the spatial model is derived) are unable to compensate for
misalignments.

%\begin{figure}
%\includegraphics[width=0.5\textwidth]{figures/wcs_rms.eps} 
%\includegraphics[width=0.5\textwidth]{figures/wcs_offset.eps} \\
%\caption{WCS}
%\label{wcs_offsets}
%\end{figure}

@@

We perturbed the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4) pixels, and to randomize the effects we
multiplied each offset by a number pulled from a normal (0,1)
distribution.  The output RMS reported by RegisterTask was noted to
track these offsets.

@@

We offset the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0) pixels.
Post--registration, this will offset the positions of sources in the
two images by the desired amount.

@@

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its Psf
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestiamte, the
postfiltered varaince is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to differences in pixel
covariance when undergoing multiple convolutions, which are not
currently tracked.  We have not yet definitively demonstrated that
this is the case here.

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/wcs_rms.eps} \\
\caption{Wcs rms.}
\label{wcsrms}
\end{figure}

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/wcs_shift.eps} \\
\caption{Wcs shift.}
\label{wcsshift}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single {\tt KernelCandidate} local
  difference image (visit=6866603, doPreFilter=True).  Columns
  indicate the Gaussian sigma of the basis, which is designed to have
  one Gaussian with the prescribed width, modified to fourth order.
  Rows indicate the amplitude of the astrometric offset that was
  introduced into the Source positions before image-to-image
  registration.  In the {\it top} row, we see that even with a small
  astrometric offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image.  The ability of the basis to compensate for bulk astrometric
  offsets is a function of the kernel width, and the kernel width is
  itself a function of the two input Psf FWHMs.  Thus there is a
  complicated dependence of the ability of the kernel to shift the
  centroids of stars.  Roughly, the basis may compensate for local
  astrometric residuals at a scale up to $\sqrt{\sigma_S^2 -
    \sigma_T^2}$.  }
\label{kernel_offsets}
\end{figure}

>>>>

We ran several pre--productions to tune the following: the dimensions
of the kernels and {\tt KernelCandidates}; the widths of the Gaussians
(sigGauss) based on the FWHM of the input images; and the scaling of
the Gaussians ($\beta$) in the basis sequence.
Figure~\ref{fig_galleryBad} provides a rogue's gallery of image
subtraction failures, centered on {\tt KernelCandidates}, and outlines
reasons for each failure.  Figure~\ref{fig_galleryGood} provides a
gallery showing what the results from a successful image subtraction
appear like, for postfiltering on the {\it top} and prefiltering on
the {\it bottom}.

@@

We implemented two simple perturbations of the inputs to the
image--to--image RegisterTask: we first added a random offset to each
object's (x,y)-coordinates, with an amplitude that was specified in
the Config and multiplied by a random number pulled from a normal
distribution; and we added a DC offset to the coordinates at an
amplitude specified in the Config.  These offsets, added to the Source
coordinates, will cause misalignments of the objects in the registered
images, as the registration is done assuming the specified positions
are correct.  In this way we are able to investigate how random
uncertainties and bulk astrometric offsets impact the false positive
rate.  We explicitly do {\it not} investigate spatial variation in
these offsets, using for example a pincushion distortion.  We
anticipate that this latter effect will be most important for spatial
interpolation and extrapolation of the matching kernel, yielding a
dipole residual field associated with the distortion.  However, we
start with the basics here to investigate the individual terms in such
misalignments.

@@

We perturbed the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0) pixels,
and to randomize the effects we multiplied each offset by a number
pulled from a normal (0,1) distribution.  The output RMS reported by
RegisterTask was noted to track these offsets.

Figure~\ref{wcsrms} shows how the numbers of false positives scales
with this RMS.  Note that the distribution is very flat out to 1.0
pixel (0.2'') for all images except for the deconvolution
configuration.  The number of false positives doubles after an RMS of
1.5--2.0''.  This test indicates that, for this very simplified case,
random astrometric uncertainties do not strongly impact the difference
image quality.  This may change with more realistic distributions of
stellar brightnesses, if individual objects pull the fit more than
others due to inverse variance weighting.  In this solution, with the
stars having the same brightness, no single random offset is likely to
drive the fit.

@@

We offset the coordinates of each template Source that was input to
RegisterTask with amplitudes of (0.025, 0.05, 0.075, 0.1, 0.125, 0.15,
0.175, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0) pixels.
Post--registration, this will offset the positions of sources in the
two images by the desired amount.

Figure~\ref{wcsshift} shows how the numbers of false positives scales
with this shift.  Surprisingly, the distribution is very flat out to
nearly 1.0 pixels (0.2'') for all images except for the deconvolution
configuration.  The number of false positives doubles after a shift of
1.5''.  This test indicates that, for this very simplified case, bulk
astrometric offsets of the degree expected from nightly processing
should not drive the false positive rate.

Figure~\ref{kernel_offsets} demonstrates the ability of the
Psf--matching kernel to redistribute flux to account for bulk
astrometric misalignments.  Since these offsets are common--mode
amongst all {\tt KernelCandidates}, the basis shapes themselves drive
the pipeline's ability to compensate for them.  Having a Gaussian
function whose sigma is commensurate with the amplitude of the offset
enables the software to place significant power in this basis term.  We
expect that, if these misalignments are a vector field, the ability of
the spatial kernel model to account for the spatial variation of the
misregistration will begin to drive the solution.  The take--away
from this analysis is that the Psf--matching basis appears flexible
enough to account for, individually, random scatter and bulk offsets
in the WCS solution.  This narrows the focus down to understanding how
spatial variation in the WCS may drive the difference image quality,
and what requirements need to be put both on the kernel basis and its
spatial variation.

@@

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its Psf
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestimate, the
postfiltered variance is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to differences in pixel
covariance when undergoing multiple convolutions, which are not
currently tracked.  We have not yet definitively demonstrated that
this is the case here.

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/wcs_rms.eps} \\
\caption{Fractional increase in false detections as a function of the
  amplitude of random astrometric offsets added to the coordinates of
  sources before image-to-image registration.  The {\it red}, {\it
    green}, and {\it blue} lines correspond to visits v6866601,
  v6866602, and v6866603 respectively, while the {\it dashed} and {\it
    solid} lines correspond to prefiltering and postfiltering,
  respectively.}
\label{wcsrms}
\end{figure}

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/wcs_shift.eps} \\
\caption{Fractional increase in false detections as a function of bulk
  astrometric offsets added to the coordinates of sources before
  image-to-image registration.  The {\it red}, {\it green}, and {\it
    blue} lines correspond to visits v6866601, v6866602, and v6866603
  respectively, while the {\it dashed} and {\it solid} lines
  correspond to prefiltering and postfiltering, respectively.}
\label{wcsshift}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single {\tt KernelCandidate} local
  difference image (visit=6866603, doPreFilter=True).  Columns
  indicate the Gaussian sigma of the basis, which is designed to have
  one Gaussian with the prescribed width, modified to fourth order.
  Rows indicate the amplitude of the astrometric offset that was
  introduced into the Source positions before image-to-image
  registration.  In the {\it top} row, we see that even with a small
  astrometric offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image.  The ability of the basis to compensate for bulk astrometric
  offsets is a function of the kernel width, and the kernel width is
  itself a function of the two input Psf FWHMs.  Thus there is a
  complicated dependence of the ability of the kernel to shift the
  centroids of stars.  Roughly, in our implementation, the basis may
  compensate for local astrometric residuals at a scale up to $\beta
  \times \sqrt{\sigma_S^2 - \sigma_T^2}$, which is the width of the
  largest Gaussian in our bases.}
\label{kernel_offsets}
\end{figure}
